\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2017-02-14 doc/publications/plos16/supplementary_information.tex
first pass through all result sections and abstract
changes to the supplementary_information.tex

<<echo=FALSE,eval=FALSE>>=

@

@@

\begin{table}[!ht]
\begin{minipage}{\textwidth}
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf Performance comparison between the different machine learning algorithms.}}
\begin{tabular}{|l|l|l|l|l|l|p{1cm}|}
\hline
\bf{Genome \%}                      & \bf{Method} & \bf{Error (ARI)} & \bf{Runtime} & \bf{Memory/Node} \\
\hline

@@

                                                    & ranger - R    & 0.94       & 4 1/2 mins       & 21GB (one processor) \\

@@

\multirow{3}{*}{phase3\_chr1: 2,504 x 6,450,364}    & \cursedforest\ & 0.063 (XX) & 1hr 58min        & 4GB \\
                                                    & RF - Spark ML  & 0.94       & 8hr 29min        & 8GB \\

@@

                                                    & ranger - R      & 0.95       & 28 mins          & num.threads=10\footnote{don't

@@

\multirow{3}{*}{phase3\_chr1-3: 2,504 x 19,328,051} & \cursedforest\ & 0.046 (XX) & 4hr 0min                                             & 4GB \\

@@

                                                    & ranger - R      &  0.97      & 58 mins\footnote{ but took 6 hours to load the

@@

@@

\multirow{4}{*}{phase3\_chr1-22: 2,504 x 81,047,467} & \cursedforest\ & XX (0.96) & 7hr 14min & 8GB \\
& RF - Spark ML & - & - & - \\
& LR - Spark ML & - & - & - \\ 
& kmeans - Spark ML  & - (0.82)      & 30hr 44min    & 24GB \\ 
& ranger (C++)       &        NA     &        NA     &            80 CPU -- runs out of memory (12.8GB per CPU) \\
& ranger (C++)       &        NA     &        NA     &            160
                                                        CPU -- runs
                                                        out of time at
                                                        20 hours \\
\hline
% \begin{tabular}{|p{1.2cm}|p{1cm}|p{1.5cm}+l|l|p{1.8cm}|p{1.5cm}|}
% \hline
% %\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \hline
% \bf{Data- set} & \bf{Samp- les} & \bf{Feat- ures}  & \bf{Method} & \bf{Error (ARI)} & \bf{Runtime} & \bf{Memory/ Node} \\
% \hline
% %\multirow{3}{*}{phase1\_chr22: 1,092 x 490,036} & CursedForest & 0.077 (XX)  &  &  \\
% phase1 chr22 &1,092 & 490,036 & CursedForest & 0.077 (XX)  & 9min 16sec & 1GB\footnote{\label{note10}10 nodes} \\
% &&& RF - Spark ML & 0.069 (0.89) & 14min 13sec & 1GB\footref{note10}  \\
% &&& RF - R &  0.05  & 24min  & 18GB\footref{note1}\footnote{using the \texttt{parallel} packages}\\
% &&& ranger -R &         0.05 &        1min 10sec  &          16GB  (one processor) \\
%  \hline
% phase3 chr22 & 2,504 & 1,103,548 &  &  &  &  \\
% &&& CursedForest & 0.053 & 9 min 8 sec & 2GB\footnote{\label{note20}20 nodes}\\
% &&& RF - Spark ML & 0.058 (0.92) & 1hr 1min & 2GB\footref{note20} \\
% &&& RF - R & NA  & NA & NA\footnote{long vectors ($> 2^31-1$)  are not supported in the .Fortran interface}\\
% &&&  H2O   &        0.05 &     16 hours       &   600GB \footnote{java -Xmx600g -jar h2o.jar,   h2o.init(nthreads=20)} \\
% \hline
% phase3 chr1 & 2,504 & 6,450,364 & CursedForest & 0.063 (XX) & 1hr 58min & 4GB \\
% &&& RF - Spark ML & (0.94) & 8hr 29min & 8GB \\
% &&& LR - Spark ML&  & 2hr 6min & 8GB\\ 
% &&& ranger  R   &  0.06   &     28 mins  &            num.threads = 10\footnote{don't know how to get the memory used, This
%   was done on dpas-03} \\ 
% \hline
% phase3 chr1-3 & 2,504 & 19,328,051  & CursedForest & 0.046 (XX) & 4hr 0min & 4GB\\
% &&& RF - Spark ML & - & -  & -\\
% &&& LR - Spark ML & & 14hr 48min & 8GB \\ 
% &&& ranger  R   &  0.05   &     58 mins\footnote{ but took 6 hours to load the data}  &      909GB\footnote{I think this is total memory. The job was run with
%   ntasks-per-node=160 cores-per-socket=10 on ruby }  \\ 
% \hline
% phase3 chr1-22 & 2,504 & 81,047,467 & CursedForest & XX (0.96) & 7hr 14min & 8GB \\ 
% &&& RF - Spark ML & - & - & - \\
% &&& LR - Spark ML & - & - & - \\ 
% %&&& kmeans - Spark ML & - (0.82) & 30hr 44min & 24GB \\ 
% \hline
\end{tabular}
\begin{flushleft} 
Comparison of different various machine learning libraries on different subsets of variant data 
from the 1000 Genomes Project.
Each random forest model consists of 50 trees.
\end{flushleft}
\label{table1}
%\end{adjustwidth}
\end{minipage}
\end{table}

\begin{table}[!ht]
\begin{minipage}{\textwidth}
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf Performance comparison between the different machine learning algorithms.}}
\begin{tabular}{|l|l|l|l|l|l|p{1cm}|}
\hline
\bf{Genome \%}                      & \bf{Method} & \bf{Error}  & \bf{Runtime} & \bf{Memory/Node} \\
\hline

@@

\multirow{4}{*}{phase3\_chr1-22: 2,504 x 81,047,467} & \cursedforest\ & XX (0.96) & 7hr 14min & 8GB \\
& Spark ML & - & - & - \\
& ranger (C++)       &        NA     &        NA     &            80 CPU -- runs out of memory (12.8GB per CPU) \\
& ranger (C++)       &        NA     &        NA     &            160
                                                        CPU -- runs
                                                        out of time at
                                                        20 hours \\
\hline
\end{tabular}
\begin{flushleft} 
Comparison of different various machine learning libraries on different subsets of variant data 
from the 1000 Genomes Project.
Each random forest model consists of 50 trees.
\end{flushleft}
\label{table1}
\end{adjustwidth}
\end{minipage}
\end{table}

>>>>

@@

\begin{sidewaystable}[!ht]
\begin{minipage}{\textwidth}
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf Performance comparison between the different machine learning algorithms.}}
\begin{tabular}{|l|l|l|l|l|l|p{1cm}|}
\hline
\bf{Genome \%}                      & \bf{Method} & \bf{Error (ARI)} & \bf{Runtime} & \bf{Memory/Node} \\
\hline

@@

@@

\multirow{3}{*}{phase3\_chr1: 2,504 x 6,450,364}    & \cursedforest\ & 0.06 & 1hr 58min        & 4GB \\
                                                    & RF - Spark ML  & 0.06       & 8hr 29min        & 8GB \\

@@

@@

\multirow{3}{*}{phase3\_chr1-3: 2,504 x 19,328,051} & \cursedforest\ & 0.04 & 4hr 0min                                             & 4GB \\

@@

@@

& ranger - R        &             0.027     &      03:58           &       pearcey --cpus-per-task=15 --mem 1600gb   906GB \\

@@

\multirow{4}{*}{phase3\_chr1-22: 2,504 x 81,047,467} & \cursedforest\ & 0.04 & 7hr 14min & 8GB \\
& RF - Spark ML & - & - & - \\
& LR - Spark ML & - & - & - \\ 
& kmeans - Spark ML  & 0.18      & 30hr 44min    & 24GB \\ 
& ranger (C++)       &        NA     &        NA     &            80 CPU -- runs out of memory (12.8GB per CPU) \\
& ranger (C++)       &        NA     &        NA     &            160
                                                        CPU -- runs
                                                        out of time at
                                                        20 hours \\
\hline
\end{tabular}
\begin{flushleft} 
Comparison of different various machine learning libraries on different subsets of variant data 
from the 1000 Genomes Project.
Each random forest model consists of 50 trees.
\end{flushleft}
\label{table1}
%\end{adjustwidth}
\end{minipage}
\end{sidewaystable}

\begin{sidewaystable}[!ht]
\begin{minipage}{\textwidth}
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf Performance comparison between the different machine learning algorithms.}}
\begin{tabular}{|l|l|l|l|l|l|p{1cm}|}
\hline
\bf{Genome \%}                      & \bf{Method} & \bf{Error}  & \bf{Runtime} & \bf{Memory/Node} \\
\hline

@@

\multirow{4}{*}{phase3\_chr1-22: 2,504 x 81,047,467} & \cursedforest\ & XX (0.96) & 7hr 14min & 8GB \\
& Spark ML & - & - & - \\
& ranger (C++)       &        NA     &        NA     &            80 CPU -- runs out of memory (12.8GB per CPU) \\
& ranger (C++)       &        NA     &        NA     &            160
                                                        CPU -- runs
                                                        out of time at
                                                        20 hours \\
\hline
\end{tabular}
\begin{flushleft} 
Comparison of different various machine learning libraries on different subsets of variant data 
from the 1000 Genomes Project.
Each random forest model consists of 50 trees.
\end{flushleft}
\label{table1}
%\end{adjustwidth}
\end{minipage}
\end{sidewaystable}

% \begin{tabular}{|p{1.2cm}|p{1cm}|p{1.5cm}+l|l|p{1.8cm}|p{1.5cm}|}
% \hline
% %\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \hline
% \bf{Data- set} & \bf{Samp- les} & \bf{Feat- ures}  & \bf{Method} & \bf{Error (ARI)} & \bf{Runtime} & \bf{Memory/ Node} \\
% \hline
% %\multirow{3}{*}{phase1\_chr22: 1,092 x 490,036} & CursedForest & 0.077 (XX)  &  &  \\
% phase1 chr22 &1,092 & 490,036 & CursedForest & 0.077 (XX)  & 9min 16sec & 1GB\footnote{\label{note10}10 nodes} \\
% &&& RF - Spark ML & 0.069 (0.89) & 14min 13sec & 1GB\footref{note10}  \\
% &&& RF - R &  0.05  & 24min  & 18GB\footref{note1}\footnote{using the \texttt{parallel} packages}\\
% &&& ranger -R &         0.05 &        1min 10sec  &          16GB  (one processor) \\
%  \hline
% phase3 chr22 & 2,504 & 1,103,548 &  &  &  &  \\
% &&& CursedForest & 0.053 & 9 min 8 sec & 2GB\footnote{\label{note20}20 nodes}\\
% &&& RF - Spark ML & 0.058 (0.92) & 1hr 1min & 2GB\footref{note20} \\
% &&& RF - R & NA  & NA & NA\footnote{long vectors ($> 2^31-1$)  are not supported in the .Fortran interface}\\
% &&&  H2O   &        0.05 &     16 hours       &   600GB \footnote{java -Xmx600g -jar h2o.jar,   h2o.init(nthreads=20)} \\
% \hline
% phase3 chr1 & 2,504 & 6,450,364 & CursedForest & 0.063 (XX) & 1hr 58min & 4GB \\
% &&& RF - Spark ML & (0.94) & 8hr 29min & 8GB \\
% &&& LR - Spark ML&  & 2hr 6min & 8GB\\ 
% &&& ranger  R   &  0.06   &     28 mins  &            num.threads = 10\footnote{don't know how to get the memory used, This
%   was done on dpas-03} \\ 
% \hline
% phase3 chr1-3 & 2,504 & 19,328,051  & CursedForest & 0.046 (XX) & 4hr 0min & 4GB\\
% &&& RF - Spark ML & - & -  & -\\
% &&& LR - Spark ML & & 14hr 48min & 8GB \\ 
% &&& ranger  R   &  0.05   &     58 mins\footnote{ but took 6 hours to load the data}  &      909GB\footnote{I think this is total memory. The job was run with
%   ntasks-per-node=160 cores-per-socket=10 on ruby }  \\ 
% \hline
% phase3 chr1-22 & 2,504 & 81,047,467 & CursedForest & XX (0.96) & 7hr 14min & 8GB \\ 
% &&& RF - Spark ML & - & - & - \\
% &&& LR - Spark ML & - & - & - \\ 
% %&&& kmeans - Spark ML & - (0.82) & 30hr 44min & 24GB \\ 
% \hline
